DATA WRANGLING
 
ABSTRACT 

The purpose of this  article is to discuss the process of data wrangling I’ve undertaken to get a clean data from the provided data. I will explain stepwise each process and give a better understanding in layman’s language .By using the R libraries , I managed to clean the data and give a better understanding of the raw data. The following is some steps I took to complete the data cleaning process.

INTRODUCTION

The following are the steps I took

Importing the data and setting our working directory : in this stage we set our working directory and  import our raw data in to our rstudio for wrangling purposes

Import libraries: here we import the specific R libraries we will to clean our data

Data exploration : here we try to get a better understanding of the data by checking on the variables and the data structures

Column renaming: as we explore the data we get long descriptive column names, since we want precise and short we rename the columns to give it a better meaning

Checking for missing values: in this step , we check for missing values in each column

Omitting the missing values: using the R command na.omit , we remove the missing values
Changing the age column in to a number : since age is a number not a character , we have to change it in to a number

Looking for outliers: we use a boxplot to check for outliers which they exist on the boxplot concerning age. We can try omit the outliers for better visualization and results

View the clean data for further cleaning: we view the cleaned data to see if we can do more. Since the data Is ready and clean we stop here

CONCLUSION

The above is the whole process I undertook during the data wrangling process. I hope you get a good understanding of each process.



---
title: "Untitled"
author: "    "
date: "7/14/2021"
output: html_document
---

importing the data we will use and setting our working directory

```{r}
setwd("C:/project 001")
library(readr)
the_counted_2015= read_csv("the-counted-2015.csv")
```

Importing libraries we will use

```{r}
library(tidyr)
library(tidyverse)
library(ggplot2)
```

data exploration

```{r}
glimpse(the_counted_2015)
names(the_counted_2015)
dim(the_counted_2015)
```

column renaming 

```{r}
names(the_counted_2015)[5]= "race"
names(the_counted_2015)[9]= "adress"
names(the_counted_2015)[12]="cause_of_death"
names(the_counted_2015)[13]="jurisdiction"
names(the_counted_2015)[1]= "id"
names(the_counted_2015)
```

checking for missing values in each column

```{r}
sum(is.na(the_counted_2015$id))
sum(is.na(the_counted_2015$name))
sum(is.na(the_counted_2015$age))
sum(is.na(the_counted_2015$gender))
sum(is.na(the_counted_2015$race))
sum(is.na(the_counted_2015$month))
sum(is.na(the_counted_2015$day))
sum(is.na(the_counted_2015$year))
sum(is.na(the_counted_2015$adress))
sum(is.na(the_counted_2015$city))
sum(is.na(the_counted_2015$state))
sum(is.na(the_counted_2015$cause_of_death))
sum(is.na(the_counted_2015$jurisdiction))
sum(is.na(the_counted_2015$armed))
```

ommiting the missing values

```{r}
my_data_new=na.omit(the_counted_2015)
names(my_data_new)
```

converting age to number since its a character which isn't correct

```{r}
my_data_new= my_data_new %>% mutate(age=as.numeric(age))
str(my_data_new)
```

looking for  outliers using a boxplot

```{r}
boxplot(my_data_new$age)

```
 
 viewing the clean data
 
```{r}
glimpse(my_data_new)
```
 
 no more action needed
 
 
 
 
 
 
 



